Mixed language processing increases cross-language phonetic transfer in Bengali-English bilinguals

### 2022-11-07 draft_BLC/BLC_AM-ID_manuscript.tex
reduced word count: discussion

Word count (5.11.2022): 9089 \\

@@

The aim of this study was to examine short-term phonetic transfer in L2, towards understanding language processing in proficient bilinguals. Two interrelated questions, discussed in section \ref{introduction}, bear on this: (i) what causes transfer in mixed-language speech; and (ii) what causes observed asymmetries between transfer in L1 vs L2.

@@

In this study, we presented nonswitched English and code-switched Bengali-English stimuli in the same session and found consistent differences between switched and nonswitched tokens. This confirms that the local act of switching does trigger transfer, even in a bilingual language mode, paralleling results from \cite{olson2016role, tsui2019impact, piccinini2015voice}. These studies focused on a temporal feature, VOT. Our results indicate that the same happens in vowels. The exact effect of transfer is interesting: since VOT is a one-dimensional measure, observing a change in VOT in a mixed language condition makes it difficult to infer whether a consonant has moved towards a specific L1/L2 counterpart, or reflects a more general system-wide shift in articulatory norms. Vowel quality measures, showing both directionality and magnitude, facilitate this. We find that productions of English vowels do not merely shift towards a Bengali counterpart, but rather towards a set of Bengali-like \textit{norms}: a vowel category that doesn't exist in Bengali nonetheless moves towards a related (perceptually close) Bengali category. This parallels results from \cite{simonet2014phonetic} for Catalan and Spanish (they examined language mode, without code-switching). The \cite{simonet2014phonetic} study showed that co-activating multiple languages within a discourse causes a transient increase in cross-language phonetic interaction, leading to changes in both L1 and L2. This is interpreted as a follows: a bilingual mode temporarily co-activates both language systems, making both L1 and L2 representations available during processing, whose interactions lead to cross-language transfer. But what does this interaction look like, and how are these abstract representations processed during speech? Examining transfer during switching can provide insights into this.

The time course of such shifts is telling: we see a difference between conditions right from the start of articulation, suggesting different acoustic targets. This suggests that such interaction involves the speech planning stage. The pattern of transfer observed in this study, particularly for \nt{2}, also demonstrates that beyond a simple L2-L1 mapping, transfer during online processing makes reference to entities beyond the level of the individual sound, demonstrating system-level interaction: a sound that does not exist in the L1 repertoire nonetheless acquires a different target, and the target is related to the L1 system in a meaningful way. This cannot be reasonably explained as articulatory inertia carried over from uttering L1 sounds. It is also difficult to account for in a purely exemplar framework, as the discourse does not contain any instances of ``L1 \nt{2}" that could affect the L2 category. If representations are combined in creative ways during online speech processing within an utterance, then can we think of intermediate representations not just for individual sounds, but rather for sound systems? While more detailed empirical data is needed to answer such questions, this approach aligns with sociophonetic understandings of bilingual phonetic variation, for example as discussed by \cite{khattab2009phonetic,khattab2013phonetic} based on the use of phonological variables by Arabic-English bilingual children. They visualize the bilingual phonetic system as ``a rich store of phonological representations", comprising of both language-internal and cross-language variation, from which the bilingual speaker is constantly choosing, guided by a variety of pressures. We propose that vowels are a good feature to examine the plausibility of such systems as cognitive models for language processing: in addition to continuous measures (unlike phonological variables), they also provide two-dimensional data (unlike VOT) that is suited to further test and refine predictions from such a model.

\paragraph{Differences between L1 vs L2 transfer:}
Previous studies of language-switching have often reported phonetic transfer from L2 to L1, but a lack of parallel transfer from L1 to L2. One postulated model that is consistent with this finding is the Inhibitory Control Model (ICM) \cite{green1986control,olson2013bilingual,tsui2019impact}. This proposes that the lack of transfer effects on L2 reflects the lower inhibition required on L2 representations in the first place, which makes switching into L2 easier, compared to the converse situation, when the high initial inhibition on L1 induces phonetic transfer while switching back into L1.  As discussed in section \ref{causes}, vowels provide a fertile testing ground for this theory at the level of sounds, as they avoid the confound of independent language-specific differences between long- and short-lag VOT languages. Both vowels in this study showed systematic shifts in the mixed-language condition compared to baseline nonswitched productions. In the context of an ICM-based analysis of between-language asymmetries in phonetic transfer as proposed by \cite{olson2013bilingual}, this is a significant finding: it demonstrates that the postulated lower switch cost on an L2 does not entirely mask transfer effects on L2 vowels. There are a few possible explanations for this. The present study focuses solely on L1 (Bengali) to L2 (English) transfer, and not the converse. It is theoretically possible that these participants are L2-dominant. That would mean that switching from L1 into L2 induces a greater switch cost, leading to transfer effects on switched L2 productions. This would lead us to expect that if the same participants are asked to switch from English to Bengali, we would not see any evidence of transfer. While more empirical data could clarify this, the language background of the present participant pool makes this scenario unlikely. All participants report regularly using both English and Bengali in a variety of settings, and feeling equally comfortable using both languages. Listening and speaking skills are expected to be most relevant to phonetic transfer in daily language use, and participants' self-ratings for these skills are at ceiling for both languages. One point to note is that participants on average rated their reading proficiency in Bengali to be lower than that in English. Given that the stimuli in this task were presented visually, differences in reading proficiency are a potential source of explanation. However, anecdotal observation suggests that phonetic transfer in L2 productions is common in daily language use too, even in the absence of reading. This raises questions about an ICM-based analysis of cross-language asymmetries in phonetic transfer that future research should investigate.

Moreover, our findings differ from \cite{tsui2019impact}, who report on the basis of VOT measurements that balanced bilinguals do not evidence an effect of language-switching. \citeauthor{tsui2019impact} attribute the absence of transfer effects to greater experience with language switching in these participants, leading to lower switch-costs in both languages. As noted in sec. \ref{participants}, the participants in this study are highly proficient bilinguals with extensive experience of language switching. Moreover, they live in an environment where they are constantly exposed to mixed-language stimuli: we expected that they are habituated to drawing on multiple language systems during online language processing. They nonetheless demonstrate cross-language transfer effects in L2 vowel production. One way to think about these differences is that vowels are generally more prone phonetic transfer than temporal features such as VOT. Thus, following from the discussion in sec. \ref{asymmetries}, it might be fruitful for future studies to focus on vowels as a potential site for transfer. This would both expand the empirical scope of existing research (see sec. \ref{asymmetry between sounds}) and avoid confounding factors arising from independent differences between long- and short-lag VOT languages. More generally, these findings point to a need for expanding the range of features in which transfer is studied, in order to make robust generalizations about the mechanisms underlying observed patterns. This is an example where studying different features of speech (e.g. VOT vs vowel quality) allows us to compare the predictions of different underlying process models. Transfer studies usually focus on a single feature. Measuring multiple features within a single design, with the same group of participants, promises to be a powerful tool to compare different models at little added cost to researchers. This would sidestep the unavoidable confounds that come with comparing results across studies, or even replication.

Beyond methodological considerations, we argue that these differences ultimately raise a broader question about a purely inhibition-based understanding of transfer. In spite of a compelling body of literature on the sociophonetic and communicative implications of phonetic variation along the L1-L2 continuum \citep[e.g.][]{khattab2013phonetic,agnihotri1979processes,stuart2011hybridity}, is it wise to think of transfer as resulting from an inability to suppress and separate language systems? Given that young children are able to use minute phonetic variation in sophisticated ways, it might be more intuitive to think of transfer as a communication-enabling feature of language, even at a purely cognitive level. Research on the cognitive mechanisms underlying cross-language interaction and variation in bilingual speech is still an emerging and exciting line of inquiry. We believe that a communication-centered approach to bilingual variation is essential to understand how bilingual language processing functions in the real world. To this end, highly proficient native multilingual populations serve as an important source of data on language behavior.

@@

Our results indicate differing effects of experimental conditions across the two vowels in both magnitude and direction of shift. Overall, shift is greater in \nt{2}. Recall (figure \ref{vowels_ie}) that the category \nt{\ae} is in a denser part of the IE vowel space than \nt{2}. Specifically: \nt{\ae} is already at the lowermost part of the vowel space. Moreover, large movements upwards in the F1 (vowel height) dimension would risk a conflation of the category with the neighboring \nt{E}. In contrast, \nt{2} (indicated as \nt{@} in the chart) is in the central part of the vowel space and is not flanked by any immediate neighboring categories. Thus, the difference in magnitude of transfer aligns with our hypothesis \textbf{H2---Asymmetry}, that the extent of shift in individual categories is constrained by the linguistic principle of contrast preservation. An alternative way to interpret this result is that the initial distance between IE \nt{2} and its target L1 category \nt{a:} was greater than that between the IE and Bengali \nt{\ae} categories, leading to a greater visible shift in the former. However, our results find parallels in previous studies with VOT that have consistently observed greater shifts in the long-lag VOT language (which has a greater acceptable range of VOT) compared to the short-lag language \citep[e.g.][]{antoniou2011inter, bullock2009trying}. This suggests that a general principle of ``room for movement" does play a role in moderating transfer. More broadly, these results suggest that phonetic transfer (i) makes reference to system-level representations that go beyond individual sounds; and (ii) respects linguistic principles. This has implications both for our understanding of phonetic transfer as a phenomenon (e.g. which mental representations and at what level of processing are involved?) and language as a system (e.g. how have transfer phenomena shaped phonological systems in multilingual settings, where languages have co-existed for long periods of time? What are the linguistic implications of increasingly cosmopolitan urban spaces and identities?).

Our results also indicate asymmetry in the direction of shift: the category \nt{2} is lowered and fronted in the mixed language condition, while \nt{\ae} is fronted. On the basis of previous literature indicating that both long-term and short-term shifts in phonological categories of multilingual speakers are targeted with respect to the other language, we hypothesized that participants' IE \nt{\ae} would shift towards the Bengali \nt{\ae}, and \nt{2} towards the Bengali \nt{a:}. As figure \ref{vowels_e_b} indicates, comparing transfer patterns with Bengali categories from \cite{shruticorpus} suggests that this is borne out.

@@

\cite{olson2013bilingual} posits that in both scripted and spontaneous code-switching paradigms, ``discursive properties may partially drive the production of code-switches, masking underlying effects of interaction", proposing cued picture-naming as an alternative. We tested whether this might independently affect the outcome of transfer, by comparing productions from the same participants in these two tasks. We found that the direction of transfer did not change across paradigms. However, there was a difference in the magnitude of transfer: in the F1 dimension, the extent of phonetic transfer is greater in the code-switching task compared to the cued picture-naming task. This asymmetry is not seen in the F2 dimension.
One way to account for this is that during code-switching, there is a greater degree of language co-activation because switches are being made within an utterance, where planning is required: Since participants see the entire utterance and then read it aloud, there is room for utterance planning. Compared to this, in the cued picture-naming paradigm participants see the language cue immediately prior the target word, inducing extremely ``local" switches and thus a smaller degree of language co-activation. While both of these are local switches and take place in a bilingual language mode, it is possible that this minute difference in time-frame induces slightly different degrees of cross-language phonetic interaction. This aligns with the findings of \cite{bullock2009trying}, who found evidence of transfer in pre-switch tokens, suggesting planning effects. In our study specifically, orthography was another factor that could have potentially affected this. In both tasks, Bengali and English words were presented in their respective orthographies. In the picture-naming task, participants saw one word at a time. The CS task in contrast presented mixed utterances in mixed orthography. It is possible that seeing mixed orthography, particularly for participants accustomed to processing mixed-language visual input in their surroundings, triggered increased language co-activation, leading to greater cross-language interaction.

We do not find any differences in the pattern of transfer, or the relative behaviors of the two categories. Our study used a frame sentence to induce code switching, and not all the items led to pragmatically meaningful sentences. Thus, one possible explanation for the overall lack of asymmetry is that the discursive effects discussed by \cite{olson2013bilingual} are triggered by more naturalistic paradigms where attention to meaning and interlocutor are important. More within-study comparisons with less structured paradigms, and spontaneous code-switching data, can clarify this. The results of the present study indicate that for the purposes of observing dynamic phonetic transfer in an experimental setup, elicitation through cued-picture-naming vs. (controlled) code-switching does not significantly alter the behavior under study. However, elicitation through code-switching allows for more planning, potentially increasing the magnitude of observable effects.

An unexpected result in this study is that the paradigm directly affected vowel quality: all vowels were more backed and lowered in the code-switching task. There is no principled reason to expect this. A closer look at the dynamics of this effect indicate that it is greatest at the beginning of articulation, and decreases steadily with time. Given that it is localized at the beginning of the target vowel, it seems likely that this lowering is an artifact of the particular stimuli used here--- resulting from some corarticulatory effect with the preceding word in the carrier phrase that was absent in the picture-naming task. While intriguing, a more detailed investigation of this effect is beyond the scope of the present study.

Although the findings of this study are interesting, we acknowledge that it has a number of limitations, including (i) a small sample size; (ii) a frame sentence for eliciting code-switching such that not all the sentences were pragmatically well-formed. This could have blocked some of the discursive effects that might result from speech planning in ordinary connected speech, thus obscuring differences between the two paradigms under study; (iii) focus on a single acoustic feature; (iv) the absence of unilingual Bengali data from the same set of speakers as the participants of the study, which would have allowed us to compare each individual's transfer patterns to their own Bengali productions, rather than to average Bengali vowel formants. Future work should address these. Our results also raise a number of open questions for future research, particularly concerning the paradigm differences and between-sound asymmetries in phonetic transfer.

\section{Conclusion}
This study investigated how the phonetic systems of highly proficient bilingual speakers interact during online speech processing, by examining the phonetic outcome of mixed-language production in speakers of Bengali and Indian English. Specifically, we examined the spectral quality of two L2 vowel categories \nt{2} and \nt{\ae} produced in two experimental conditions: nonswitch (English) and switch/mixed (switching from Bengali to English).
Comparisons of F1 and F2 across conditions reveal that mixed language processing temporarily increases phonetic interaction between the languages, causing the English \nt{\ae} of participants to converge towards the Bengali category \nt{\ae}, and \nt{2} to converge towards a related L1 category \nt{a:}. The extent and direction of shift is asymmetrical across the two categories, which parallels findings in prior studies. This asymmetry appears to be mediated by the language-specific distribution of vowel categories. These findings evidence an effect of phonetic transfer on L2 vowels, and raise questions for proposed inhibition-based accounts of cross-language phonetic interaction. Mixed productions were elicited in two switching paradigms: code-switching and cued picture-naming, to verify whether the postulated discursive elements in code-switching independently influence the outcome of phonetic transfer. Results suggest that in an experimental setup, elicitation through cued-picture-naming vs. (controlled) code-switching does not significantly alter the behavior under study. However, the magnitude of observed transfer is greater in the code-switching paradigm. These findings add to empirical data, highlight the need for diversifying the set of language pairs and populations in transfer studies, and encourage more within-study comparisons of paradigms and features, towards establishing better methodological norms. Ultimately, these results contribute to ongoing discussion about the implications of short-term phonetic interaction as a window to understanding language as a mental and social system.

>>>>

Word count: \\

@@

This study examined short-term phonetic transfer in L2, towards understanding language processing in proficient bilinguals. Two interrelated questions (section \ref{introduction}), bear on this: (i) what causes transfer in mixed-language speech; (ii) what causes observed asymmetries between transfer in L1 vs L2.

@@

We presented nonswitched English and code-switched Bengali-English stimuli in the same test block and found consistent differences in vowel quality. This confirms that the local act of switching triggers transfer, even in a bilingual language mode, paralleling results from \citet{olson2016role, tsui2019impact, piccinini2015voice}. These studies focused on a temporal feature, VOT. Our results demonstrate a similar effect in vowels. Since VOT is a one-dimensional measure, a difference across conditions is insufficient to infer whether the category has shifted towards a specific L1/L2 counterpart, or reflects a more general system-wide shift in articulatory norms. Vowel quality measures, having both directionality and magnitude, enable this. We find that productions of English vowels do not merely shift towards a Bengali counterpart, but rather towards a set of Bengali-like \textit{norms}: a vowel category that doesn't exist in Bengali nonetheless moves towards a related (perceptually close) Bengali category. This parallels findings from \cite{simonet2014phonetic} for Catalan and Spanish (they manipulated language mode, without code-switching). The \cite{simonet2014phonetic} study showed that co-activating multiple languages within a discourse causes a transient increase in cross-language phonetic interaction, leading to changes in both L1 and L2. This is interpreted as a follows: a bilingual mode temporarily co-activates both language systems, making both L1 and L2 representations available during processing, whose interactions lead to cross-language transfer. But what does this interaction look like, and how are these abstract representations processed during speech? Examining transfer during switching can provide insights into this.

The time course of such shifts is telling: we see a difference between conditions right from the start of articulation, suggesting different acoustic targets. This indicates that such interaction involves the speech planning stage. The pattern of transfer observed in this study, particularly for \nt{2}, also demonstrates that beyond a simple L2-L1 mapping, transfer during online processing makes reference to entities above the level of individual sounds, demonstrating system-level interaction: a sound that does not exist in the L1 repertoire nonetheless acquires a different target, and the target is related to the L1 system in a meaningful way. This cannot be reasonably explained as articulatory inertia carried over from uttering L1 sounds. It is also difficult to account for in a purely exemplar framework, as the discourse does not contain any instances of ``L1 \nt{2}" that could affect the L2 category. If representations are combined in creative ways during online speech processing, then can we think of intermediate representations not just for individual sounds, but rather for sound systems? While more detailed empirical data is needed to answer such questions, this approach aligns with sociophonetic understandings of bilingual phonetic variation, e.g. as discussed by \cite{khattab2009phonetic,khattab2013phonetic} based on the use of phonological variables by Arabic-English bilingual children. These visualize the bilingual phonetic system as ``a rich store of phonological representations", comprising of both language-internal and cross-language variation, from which the bilingual speaker is constantly choosing, guided by a variety of pressures. We propose that vowels are a good feature to examine the plausibility of such systems as cognitive models for language processing: in addition to continuous measures (unlike phonological variables), they also provide two-dimensional data (unlike VOT) that is suited to further test and refine predictions from such a model.

\paragraph{Differences between L1 vs L2 transfer:}
Existing studies on language-switching have often reported phonetic transfer from L2 to L1, no parallel transfer from L1 to L2. One postulated model consistent with this finding is the Inhibitory Control Model (ICM) \cite{green1986control,olson2013bilingual,tsui2019impact}. %This proposes that the lack of transfer effects on L2 reflects the lower initial inhibition required on L2 representations, which makes switching into L2 easier, compared to the converse situation, where the high initial inhibition on L1 induces phonetic transfer while switching back into L1.
As discussed in section \ref{causes}, vowels provide a fertile testing ground for this theory at the level of sounds. Both vowels in this study showed systematic shifts in the mixed-language condition compared to baseline nonswitched productions. In the context of an ICM-based analysis of between-language asymmetries in phonetic transfer, as proposed by \cite{olson2013bilingual}, this demonstrates that the postulated lower switch cost on L2 does not entirely mask transfer effects on L2 vowels. There are a few possible explanations for this. The present study focuses solely on L1 (Bengali) to L2 (English) transfer, and not the converse. It is theoretically possible that these participants are L2-dominant. That would mean that switching from L1 into L2 induces a greater switch cost, leading to transfer effects on switched L2 productions. We should then expect a lack of transfer while switching from English to Bengali. While more empirical data could clarify this, the language background of these participants makes this scenario unlikely. All individuals report regularly using both English and Bengali in a variety of settings, and feeling equally comfortable using both languages. Listening and speaking skills are expected to be most relevant to phonetic transfer in daily language use, and participants' self-ratings for these skills are at ceiling for both languages. One point to note is that participants on average rated their reading proficiency in Bengali to be lower than in English. Given that the stimuli in this task were presented visually, differences in reading proficiency are a potential source of explanation. However, anecdotal observation suggests that phonetic transfer in L2 productions is common in daily language use too, even in the absence of reading. This raises questions about an ICM-based analysis of cross-language asymmetries in phonetic transfer that future research should investigate.

Moreover, our findings differ from \cite{tsui2019impact}, who report on the basis of VOT measurements that balanced bilinguals do not evidence an effect of language-switching. \citeauthor{tsui2019impact} attribute this to greater language-switching experience in these participants, leading to lower switch-costs in both languages. As noted in sec. \ref{participants}, participants in this study are highly proficient bilinguals with extensive experience of language switching. Moreover, their environment constantly exposes them to mixed-language stimuli: we expect that they are habituated to drawing on multiple language systems during online language processing. They nonetheless demonstrate cross-language transfer effects in L2 vowel production. There are a few different ways to approach this:

One possible explanation is that vowels are generally more susceptible to phonetic transfer than temporal features such as VOT. Following from the discussion in sec. \ref{asymmetries}, it might therefore be fruitful for future studies to focus on vowels as the site for transfer. More generally, these findings point to a need for expanding the range of features studied, in order to make robust generalizations about the mechanisms underlying observed patterns of transfer. This is an example where examining different features of speech (e.g. VOT vs. vowel quality) allows us to compare predictions from different underlying process models. Measuring multiple features within a single study promises to be a powerful tool for interpretation and theorizing, at little added cost to researchers. This would sidestep the unavoidable confounds that come with comparing results across studies.

Beyond methodological considerations, however, we argue that these differences ultimately raise broader questions about a purely inhibition-based understanding of transfer. Given the compelling body of literature on the socio-communicative significance of phonetic variation along the L1-L2 continuum \citep[e.g.][]{khattab2013phonetic,agnihotri1979processes,stuart2011hybridity}, is it wise to think of transfer as resulting from an inability to suppress and separate language systems? Seeing as even young children are capable of using minute phonetic variation in sophisticated ways, it might be more intuitive to think of transfer as a communication-enabling feature of language, even at a purely cognitive level. Research on the cognitive mechanisms underlying cross-language interaction and variation in bilingual speech is still an emerging and exciting line of inquiry. We believe that a communication-centered approach to bilingual variation is essential to understand how bilingual language processing functions in the real world. To this end, highly proficient native multilingual populations serve as an important source of data.

@@

Our results indicate differences across the two vowels in both magnitude and direction of shift. Overall, shift is greater in \nt{2}. Recall (figure \ref{vowels_ie}) that the category \nt{\ae} is in a denser part of the IE vowel space than \nt{2}. Specifically: \nt{\ae} is already at the lowermost part of the vowel space. Large movements upwards in the F1 (vowel height) dimension would risk a conflation of the category with the neighboring \nt{E}. In contrast, \nt{2} (indicated as \nt{@} in the chart) is in the central part of the vowel space and not flanked by any immediate neighboring categories. Thus, the difference in magnitude of transfer aligns with our hypothesis \textbf{H2---Asymmetry}, that the extent of shift in individual categories is constrained by the linguistic principle of contrast preservation. An alternative way to interpret this result is that the initial distance between IE \nt{2} and its target L1 category \nt{a:} was greater than that between the IE and Bengali \nt{\ae} categories, leading to a greater visible shift in the former. However, our results find parallels in previous studies with VOT that have consistently observed greater shifts in the long-lag VOT language (greater acceptable range of VOT) compared to the short-lag language \citep[e.g.][]{antoniou2011inter, bullock2009trying}. This suggests that a general principle of ``room for movement" does play a role in moderating transfer. These results suggest that phonetic transfer (i) makes reference to system-level representations that go beyond individual sounds; and (ii) respects linguistic principles. This has implications both for our understanding of phonetic transfer as a phenomenon (e.g. which mental representations and at what level of processing are involved?) and language as a system (e.g. how have transfer phenomena shaped phonological systems in multilingual settings, where languages have co-existed for long periods of time? What are the linguistic implications of increasingly cosmopolitan urban spaces and identities?).

Our results also indicate asymmetry in the direction of shift: the category \nt{2} is lowered and fronted in the mixed language condition, while \nt{\ae} is fronted. On the basis of previous literature indicating that both long-term and short-term shifts in categories of multilingual speakers are targeted with respect to the other language, we hypothesized that participants' IE \nt{\ae} would shift towards the Bengali \nt{\ae}, and \nt{2} towards the Bengali \nt{a:}. As figure \ref{vowels_e_b} indicates, comparing transfer patterns with Bengali categories from \cite{shruticorpus} suggests that this is borne out.

@@

\cite{olson2013bilingual} posits that in both scripted and spontaneous code-switching, ``discursive properties may partially drive the production of code-switches, masking underlying effects of interaction", proposing cued picture-naming as an alternative. We tested whether this might independently affect the outcome of transfer, by comparing productions from the same participants in these two tasks. The direction of transfer did not change across paradigms. However, there was a difference in the magnitude of transfer: in the F1 dimension, the extent of phonetic transfer is greater in the code-switching task compared to the cued picture-naming task. This asymmetry is not seen in the F2 dimension.
It is possible that code-switching induces a greater degree of language co-activation because switches are being made within an utterance, which requires planning: participants see the entire sentence and then read it aloud. Compared to this, the cued picture-naming paradigm presents the language cue immediately prior to the target word, inducing extremely ``local" switches and thus a smaller degree of language co-activation. While both of these are local switches and take place in a bilingual language mode, it is possible that this minute difference in time-frame induces slightly different degrees of cross-language phonetic interaction. This aligns with the findings of \cite{bullock2009trying}, who found evidence of transfer in pre-switch tokens, suggesting planning effects. In our study specifically, orthography was another factor that might have affected this. In the picture-naming task, participants saw one word at a time. The CS task in contrast presented mixed utterances in mixed orthography. It is possible that seeing mixed orthography, particularly for individuals accustomed to processing mixed-language visual input in their surroundings, triggered increased language co-activation, leading to greater cross-language interaction.

We do not find any differences in the pattern of transfer across paradigms. Our study used a frame sentence to induce code switching, and not all the items led to pragmatically meaningful sentences. Thus, one possible explanation for the overall lack of asymmetry is that the discursive effects discussed by \cite{olson2013bilingual} are triggered by more naturalistic paradigms where attention to meaning and interlocutor are important. More within-study comparisons with less structured paradigms, and spontaneous code-switching data, can clarify this. The results of this study indicate that for the purposes of observing dynamic phonetic transfer in an experimental setup, elicitation through cued-picture-naming vs. (controlled) code-switching does not significantly alter the behavior under study. However, elicitation through code-switching allows for more planning, potentially increasing the magnitude of observable effects.

An unexpected result in this study is that the paradigm directly affected vowel quality: all vowels were more backed and lowered in the code-switching task. There is no principled reason to expect this. A closer look at the dynamics of this effect indicate that it is greatest at the beginning of articulation, and decreases with time. Given that it is localized at the beginning of the target vowel, it likely an artifact of the particular stimuli used here. While intriguing, a more detailed investigation of this effect is beyond the scope of the present study.

Although the findings of this study are interesting, we acknowledge that it has a number of limitations, including (i) a small sample size; (ii) a frame sentence for eliciting code-switching such that not all the sentences were pragmatically well-formed. This could have blocked some of the discursive effects that might result from speech planning in ordinary connected speech, thus obscuring differences between the two paradigms under study; (iii) focus on a single acoustic feature; (iv) the absence of unilingual Bengali data from the same set of speakers as the participants of the study, which would have allowed us to compare each individual's transfer patterns to their own Bengali productions. Future work should address these. Our results also raise a number of open questions for future research, particularly concerning paradigm differences and between-sound asymmetries in phonetic transfer.

\section{Conclusion}
This study investigated how the sound systems of highly proficient bilingual speakers interact during online speech processing, by investigating the phonetic outcome of mixed-language production in speakers of Bengali and Indian English. We examined the spectral quality of L2 vowel categories \nt{2} and \nt{\ae} produced in two experimental conditions: nonswitch (English) and switch/mixed (switching from Bengali to English).
Comparisons of F1 and F2 across conditions reveal that mixed language processing temporarily increases phonetic interaction between the languages, causing the English \nt{\ae} of participants to converge towards the Bengali category \nt{\ae}, and \nt{2} to converge towards a related L1 category \nt{a:}. The extent and direction of shift is asymmetrical across the two categories, which parallels findings in prior studies. This asymmetry appears to be mediated by the language-specific distribution of vowel categories. These findings evidence an effect of phonetic transfer on L2 vowels, and raise questions for proposed inhibition-based accounts of cross-language phonetic interaction. Mixed productions were elicited in two switching paradigms: code-switching and cued picture-naming, to verify whether the postulated discursive elements in code-switching independently influence the outcome of phonetic transfer. Results suggest that in an experimental setup, paradigm does not significantly alter the behavior under study. However, the magnitude of observed transfer is greater in the code-switching paradigm. These findings add to empirical data, highlight the need for diversifying the set of language pairs and populations in transfer studies, and encourage more within-study comparisons of paradigms and acoustic features, towards establishing better methodological norms. Ultimately, these results contribute to ongoing discussion about the implications of short-term phonetic interaction as a window to understanding language as a mental and social system.

### 2022-11-07 draft_BLC/extra_text_from_manuscript/extra.tex
reduced word count
reduced word count: discussion

@@

>>>>

The significant Time*Vowel interaction shows that the F1 trajectory differs between the vowels as expected, since the two categories have different acoustic targets.

@@

The exact nature of transfer is interesting: since VOT is a one-dimensional measure, observing a change in VOT in a mixed language condition makes it difficult to infer whether a consonant has moved towards a specific L1/L2 counterpart, or reflects a more general system-wide shift in articulatory norms. Vowel quality measures, showing both directionality and magnitude, facilitate this. We find that productions of English vowels do not merely shift towards a Bengali counterpart, but rather towards a set of Bengali-like \textit{norms}: a vowel category that doesn't exist in Bengali nonetheless moves towards a related (perceptually close) Bengali category. This parallels results from \cite{simonet2014phonetic} for Catalan and Spanish (they examined language mode, without code-switching). The \cite{simonet2014phonetic} study showed that co-activating multiple languages within a discourse causes a transient increase in cross-language phonetic interaction, leading to changes in both L1 and L2. This is interpreted as a follows: a bilingual mode temporarily co-activates both language systems, making both L1 and L2 representations available during processing, whose interactions lead to cross-language transfer. But what does this interaction look like, and how are these abstract representations processed during speech? Examining transfer during switching can provide insights into this.

This proposes that the lack of transfer effects on L2 reflects the lower initial inhibition required on L2 representations, which makes switching into L2 easier, compared to the converse situation, where the high initial inhibition on L1 induces phonetic transfer while switching back into L1.
